DRAFT: Direct-style process creation on Linux

### 2020-05-15 research/paper/paper.tex
paper

On most systems today,
the interfaces for process creation
can be divided into "fork-style" and "spawn-style".
A few systems use a third style, which we refer to as "direct-style".
\subsection{Fork-style}
By far the most well-known fork-style interface is Unix's fork.\cite{forkhist}
With a fork-style interface,
a process calls \texttt{fork}, and a copy of that process is made,
with the return value of fork being the only differentiation between the process and its copy.
The same program continues executing in both processes.
Typically, the program will case on the return value to determine if it's running in the parent or the child,
and run process setup and initialization code if the program is running in the child,
calling various syscalls to set up the state of the child.

@@

In direct-style, a process is created by its parent in a non-running state,
and then supplied with various resources,
and then started running once it is fully set up.
In operating systems with true direct-style process creation,
the syscalls that can mutate a process
take explicit arguments to indicate which process they should operate on.
In this way, the same syscalls that can mutate a process while it is running,
are called by the parent process to mutate the process while it is being set up.

>>>>

On most systems today,
the interfaces for process creation
can be divided into "fork-style" and "spawn-style".
A few systems use a third style, which we refer to as "direct-style".
\subsection{Fork-style}
By far the most well-known fork-style interface is Unix's fork.\cite{forkhist}
With a fork-style interface,
a process, called the parent process, calls \texttt{fork},
and a copy of that process is made,
called the child process,
differing only in a few attributes such as process id.
The same program continues executing in both processes,
with a different return value from fork to allow it to distinguish whether it is running in the parent or the child.
Typically, the program will case on the return value
and, if running in the child,
call various system calls to modify the state of the new process.

@@

In direct-style, a process is created by its parent in a non-running state,
then supplied with various resources,
and then started running once it is fully set up.
In operating systems with true direct-style process creation,
the syscalls that can mutate a process
take explicit arguments to indicate the process on which they should operate.
In this way, the same syscalls that can mutate a process while it is running,
are called by the parent process to mutate the process while it is being set up.
